---
title: "Portfolio 1, Study Group 10"
author: "Jesper Fischer, Lasse Hansen & Sarah Hvid Andersen"
output:
  html_document: default
  pdf_document: default
---

```{r}
#Loading packages
pacman::p_load(sjPlot, tidyverse, broom)
```

#### 1a

```{r}
#importing the data.
sleep <-  read.csv("sleepstudy.csv")
#selecting the data from subject 308
p308 <- subset(sleep, Subject == 308)
#making a model with the data from subject 308
m1 <-  lm(Reaction ~ Days, data = p308)
summary(m1)
```
Days participant 308 was sleep deprived explained a significant proportion of variance in reaction time, R^2 = .64, F(1, 8) = 17.12, p = .003.

#### 1b

As reported above the model degrees of freedom = 1 and the residual degrees of freedom = 8.


#### 1c + 1d

```{r}
dist_f(f=0, deg.f1 = 1, deg.f2 = 8,xmax=20)
```

A regression with these degrees of freedom becomes statistically significant when the f-value gets above 5.32 at p < .05.

#### 2a

```{r}
#getting the coefficients (slope and intercept) for each participant:
fitted_models <-  sleep %>% group_by(Subject) %>% do(model = lm(Reaction ~ Days, data = .))
#putting the data, into a tibble
m2 <-  fitted_models %>% tidy(model)
m2
```

#### 2b

```{r}
#making subject a factor:
sleep$Subject <-  as.factor(sleep$Subject)
#plotting Reaction time as a function of days sleep deprieved with each subject having their own regression line:
ggplot(sleep, aes(Days, Reaction, colour = Subject))+geom_point()+geom_smooth(method = "lm", se=FALSE) + xlab("Days of Sleep Deprivation") + ylab("Reaction Time in mS")
```


#### 2c

```{r}
#puting the degrees of freedom into the dataframe with the regression coefficients
m2$t.value <- m2$statistic
m2$statistic <- NULL
m2$df_model <-  1
m2$df_residual <-  8
m2
```

#### 2d

```{r}
#We want to look at the effects sleep deprivation has on reaction time therefore we only look at the slopes. The intercepts only show the relationship between the mean reaction time and the reaction time when the participants were 0 days sleep deprived.

m3 <-  filter(m2, term == "Days")

#We only want those that displayed a statistically significant effect from sleep deprivation:
m4 <-  filter(m3, p.value < 0.05)
m4
```

14 individuals displayed a statistically significant effect of sleep deprivation.

#### 3a + 3b
We used an one tailed t-test because we were asked wether the slopes are larger than zero. there is strong theoretical background that the slopes are positive (reationtime increases as days sleep deprived increases). Therefore it is justifyable to use a one-tailed t-test even though the risk of comitting a type 1 error is increased. A t-test was used because we want to compare two means (0 and the mean of the reaction time).
This test assumes normally distributed data, which is tested below with visual inspection and by using the shapiro wilks test. 

```{r}
#first accessing the slopes from the previous exercise
est <-  m3$estimate

#Putting them into a dataframe
estimates <-  as.data.frame(est)

#testing whether the slopes are normally distributed:
shapiro.test(est)

ggplot(estimates, aes(est)) + geom_histogram(aes(y=..density..), colour = "black", fill = "white", binwidth = 3) + labs(x ="Slope estimates", y = "Density") + stat_function(fun = dnorm, args = list(mean = mean(estimates$est), sd = sd(estimates$est)), colour= "black", size = 1)
```

As can be seen the data looks kind of normally distributed, and the shapiro wilks test is non-significant indicating that the reaction times are not significantly different from a perfect normal distribution.

One can therefore perform the t-test:

```{r}
#Performing the t-test and telling it to perform a one-sided kind
test <-  t.test(est, mu = 0, alternative = "two.sided")
test

#Calculating the effect size r
r <-  sqrt(test$statistic^2/((test$statistic^2)+test$parameter))
r
```

#### 3c

Our slopes are shown to be statistically  significantly different from 0, t(17)=6.77  p < .001, r = 0.85

#### 3d 
```{r}
#Making a plot of Reaction time as a function of Days. And adding the mean reaction time with standard error bars for each day 
ggplot(sleep, aes(x = Days, y = Reaction))+geom_point(stat = "summary", fun.y = mean)+ stat_summary(fun.data = mean_se, geom = "errorbar", color = 'black', width = 0.1)+geom_smooth(method = "lm", alpha=0) + xlab("Days of Sleep Deprivation") + ylab("Reaction Time in mS")
```


#### Voluntary bonus task:

```{r}
#Adding 10% noise.
sleep$Reaction.Noise <-  sleep$Reaction+sleep$Reaction*runif(10, min = -0.05, max = 0.05)
```


```{r}
#Getting statistics from the newly derived data
fitted_models1 <-  sleep %>% group_by(Subject) %>% do(model1 = lm(Reaction.Noise ~ Days, data = .))

#putting the data, into a tibble(dataframe)
m5 = fitted_models1 %>% tidy(model1)

#Filtering out the slopes
m6 = filter(m5, term == "Days")

#We only want those that displayed a statistically significant effect from sleep deprivation:
m7 = filter(m6, p.value < 0.05)

m7
```





```{r}
#Saving the estimates and putting them into a dataframe
est1 = m6$estimate
estimates1 = as.data.frame(est1)

#Testing whether the slopes are normally distributed:
shapiro.test(est1)

#Plotting a histogram of the new data
ggplot(estimates1, aes(est1)) + geom_histogram(aes(y=..density..), colour = "black", fill = "white", binwidth = 3) + labs(x ="Slope estimates", y = "Density") + stat_function(fun = dnorm, args = list(mean = mean(estimates1$est1), sd = sd(estimates1$est1)), colour= "black", size = 1)
```

As can be seen the data looks kind of normally distributed, and the shapiro wilks test is non-significant indicating that the reaction times are not significantly different from a perfect normal distribution.

One can therefore perform the t-test with the new data:

```{r}
#Dping a t-test and telling it to perform a one-sided type
test1 = t.test(est1, mu = 0, alternative = "two.sided")
test1

#Calculating the effect size
r1 = sqrt(test1$statistic^2/((test1$statistic^2)+test1$parameter))
r1
```

As can be seen there is only 13 statistically significant slopes after adding 10% white noise.

The results show that adding 10% white noise can have an influence on the individual level, 13 slopes being statistically significant meaning that one person after adding the 10% noise did not seem to have their reaction time influenced by sleep deprivation, eventhough they seemed to be influenced by sleep deprivation without the 10% noise. 

On a group level adding the 10% noise had no meaningful effect. 
